---
title: "Assignment"
author: "Megan Thong"
date: "02/07/2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/James/Documents/JHU, R Coursera/RepData_PeerAssessment1")
library(dplyr)
```

## Loading and preprocessing the data

The zip file of the data is already in the GitHub repo. Here, we unzip it to create a csv file:

```{r load_data}
zipFile <- "activity.zip"
data <- "Data"
if (!file.exists(data)) {
  unzip(zipFile)
}

activity = read.csv("activity.csv", header=TRUE, sep=",", na.strings="NA")
```

Only preprocessing we really need to do is transform the "date" column into an actual Date format:

```{r date}
class(data$date)
data$date = as.Date(data$date, "%Y-$m-$d")
```

## What is mean total number of steps taken per day?

In order to calculate this, we need to first exclude missing values, then sum the number of steps per day:

```{r steps}
#remove NAs
data1 = data[!is.na(data$steps), ]

#calculate total number of steps each day
group_by(data1, date) %>% summarise(total = sum(steps))
steps_total = aggregate(steps ~ date, data = data1, sum)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
